Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-13 eswc-2012/phdsymp.tex
refactor federation sota
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@93 43ef1190-90fd-4dd7-a769-d0121f02b80e

\section{Introduction}

@@

\section{Problem Definition}

@@

When using FedBench, (see \cite{Schmidt:2011:FBS:2063016.2063054}),
the authors note that the absence of statistics has very adverse
effects on performance as join order is impossible to optimize. It is
also pointed out that it causes other problems as well, for example a
high number of requests to federation members.

This problem has already received some attention, \cite{5337556}
addressed this problem by making it possible to exchange histograms
and other data. However, histogram approaches generally suffers from
the problem that they grow too large or become a insufficiently
accurate digest, especially in face of very heterogeneous data.

@@

\section{State of the Art}

\paragraph{In SPARQL Federation}

A recent scientific treatment of the current specification, is in
\cite{springerlink:10.1007/978-3-642-21064-8-1}. In this paper, the
authors also show an optimization strategy based on execution order of
so-called well-designed patterns.

In \cite{5337556}, it is shown how various statistics-gathering
techniques, including histograms can be computed to assist the SemWIQ
optimizer, written by the same authors.
\cite{Harth:2010:DSO:1772690.1772733} introduced QTrees, which may
alleviate the problem of histogram size, but it may not solve it.

For benchmarking federated query processing strategies,
\cite{Schmidt:2011:FBS:2063016.2063054} report the development of \emph{FedBench}.

>>>>

\section{Motivation}

@@

\section{State of the Art and Open Problems}

@@

A recent scientific treatment of the current specification is in
\cite{springerlink:10.1007/978-3-642-21064-8-1}. In this paper, the
authors also show an optimization strategy based on execution order of
so-called well-designed patterns.

For benchmarking federated query processing strategies,
\cite{Schmidt:2011:FBS:2063016.2063054} report the development of
\emph{FedBench}. In this paper, the authors note that the
absence of statistics has very adverse effects on performance as join
order is impossible to optimize. It is also pointed out that it causes
other problems as well, for example a high number of requests to
federation members. I take this as a key research problem.

This problem has already received some attention, in \cite{5337556} it
is shown how various statistics-gathering techniques, including
histograms can be computed to assist the SemWIQ optimizer, written by
the same authors.  However, histogram approaches generally suffers
from the problem that they grow too large or become a insufficiently
accurate digest, especially in face of very heterogeneous data.
\cite{Harth:2010:DSO:1772690.1772733} introduced QTrees, which may
alleviate the problem of histogram size, but it may not solve it.

@@
